---
title: "Building and checks"
output:
  html_document:
    toc: true
    toc_depth: 2
---

# Introduction

Building and checking are two essential steps to finalize your R package. Building means converting your package into a .tar file ("TTape ARchiver" or "tarball") which converts the folder structure of your package into a single file which can easily be read by machines and distributed. Cechking means runing a set of tests on your package, ensuring, that it follows R apckage conventions, is portable across platforms and does not cause problems or damage to the users.

# Build the package
To build your package, either use the build tab of Rstudio or run `build(".")` in the package directory.

```{r, eval = FALSE}
build(".")
```

# Check the package

Building seems to have worked! But let's make sure that everything has been set up properly.

# Tidy code and goodpractice

On top of `check~ we will run two additional test to make sure the code of your package is readable and does not cause problems. First, run from the formatR package.




# Exercises

1. Run R CMD check on your package
2. Build your package
3. Run `formatr` and goodpractice on the code

source("https://install-github.me/MangoTheCat/goodpractice")
library(goodpractice)
gp("C:/Users/az64mycy/Dropbox (iDiv)/research_projects/29_CoordinateCleaner/CoordinateCleaner")

#R CMD check
devtools::check_built(path = "", 
                      manual = T, cran = T)

devtools::spell_check()
3. commit and push to GitHub



```{r, eval=FALSE}
check()
#> ── R CMD check results ─────────────────────────────────────────────────────────── myPackage 0.0.0.9000 ────
#> Duration: 14.2s
#> 
#> ❯ checking DESCRIPTION meta-information ... WARNING
#>   Non-standard license specification:
#>     What license it uses
#>   Standardizable: FALSE
#> 
#> ❯ checking for missing documentation entries ... WARNING
#>   Undocumented code objects:
#>     ‘greet’
#>   All user-level objects in a package should have documentation entries.
#>   See chapter ‘Writing R documentation files’ in the ‘Writing R
#>   Extensions’ manual.
#> 
#> 0 errors ✔ | 2 warnings ✖ | 0 notes ✔
```