---
title: "Package skeleton"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/home/se87kuhe/my_science/r-dev")
```

# Introduction

Initial blabla about package skeleton

```{r}
library(devtools)
```


# Create the package

```{r, eval=FALSE}
create_package("/home/se87kuhe/my_science/r-dev/myPackage", rstudio = TRUE, open = TRUE)
#> ✔ Creating '/home/se87kuhe/my_science/r-dev/myPackage/'
#> ✔ Setting active project to '/home/se87kuhe/my_science/r-dev/myPackage'
#> ✔ Creating 'R/'
#> ✔ Writing 'DESCRIPTION'
#> Package: myPackage
#> Title: What the Package Does (One Line, Title Case)
#> Version: 0.0.0.9000
#> Authors@R (parsed):
#>     * First Last <first.last@example.com> [aut, cre] (<https://orcid.org/YOUR-ORCID-ID>)
#> Description: What the package does (one paragraph).
#> License: What license it uses
#> Encoding: UTF-8
#> LazyData: true
#> ✔ Writing 'NAMESPACE'
#> ✔ Writing 'myPackage.Rproj'
#> ✔ Adding '.Rproj.user' to '.gitignore'
#> ✔ Adding '^myPackage\\.Rproj$', '^\\.Rproj\\.user$' to '.Rbuildignore'
#> ✔ Opening '/home/se87kuhe/my_science/r-dev/myPackage/' in new RStudio session
#> ✔ Setting active project to '/home/se87kuhe/my_science/teaching/yDiv_building_r_packages'
```


# Set up git(hub)

The previous step will have started another instance of RStudio. There you need to call `library(devtools)` again.

```{r, eval=FALSE}
library(devtools)
use_git()
#> ✔ Setting active project to '/home/se87kuhe/my_science/r-dev/myPackage'
#> ✔ Initialising Git repo
#> ✔ Adding '.Rhistory', '.RData' to '.gitignore'
#> There are 5 uncommitted files:
#> * '.gitignore'
#> * '.Rbuildignore'
#> * 'DESCRIPTION'
#> * 'myPackage.Rproj'
#> * 'NAMESPACE'
#> Is it ok to commit them?
#> 
#> 1: Negative
#> 2: For sure
#> 3: No way
#> 
#> Selection: 2
#> ✔ Adding files
#> ✔ Commit with message 'Initial commit'
#> ● A restart of RStudio is required to activate the Git pane
#> Restart now?
#> 
#> 1: Nope
#> 2: I agree
#> 3: Absolutely not
#> 
#> Selection: 2
```

Which would restart RStudio once again to update its interface for git-functionality (i.e., add buttons to access and manage git). Call `library(devtools)` once again.

Now set up a github repository (at [www.github.com](www.github.com)), into which the code will be pushed.

# Set a license

A bit on licenses?

```{r, eval=FALSE}
use_gpl3_license(name = "me")
#> ✔ Setting License field in DESCRIPTION to 'GPL-3'
#> ✔ Writing 'LICENSE.md'
#> ✔ Adding '^LICENSE\\.md$' to '.Rbuildignore'
```


# Edit `DESCRIPTION`

which fields do exist here and what do they mean?

Some fields are gonna be added automatically with other `usethis` functions.

## Versioning

## Include your OrcID 
